Comparing variability across and within fMRI analysis software packages

### 2021-12-22 paper/paper.tex
Fix figures and refs

As a result of these factors, the variability resulting from the use of
different fMRI analysis tools implementing similar analytical approaches
(between-tool variability) can reach worrying magnitudes. For instance, the
study in~\cite{bowring2019exploring} compared the results produced by the
three main fMRI analysis toolboxes, namely SPM, AFNI and FSL, using similar
pipelines. It reported limited similarity between the activation clusters
produced by these tools, measured by Dice coefficients ranging from 0.0 to
0.769, where 0 indicates non-overlapping
clusters and 1 means identical clusters. More recently, the work
in~\cite{Li2021.12.01.470790} also showed a low similarity between the
results produced by five different tools (ABCD, CCS, CPAC, DPARSF, fMRIPrep
\tristan{Ali, could you add refs to these tools from Li et al 2021? Also
for SPM, FSL and AFNI in the previous sentence}) and identified the main
factors contributing to these differences. The magnitude of the highlighted
differences suggest that between-tool variability may be playing a critical
role in the reproducibility of fMRI analysis overall.

@@

We evaluated BT variability for thresholded as well as unthresholded
group-level and subject-level t-statistics maps by computing the absolute
differences of t-statistic maps across tools. For WT, we computed the
average absolute difference across the three Fuzzy Libmath samples.
Moreover, we computed WT variability for a pair of tools as the sum of WT values for each tool.
\tristan{Ali, can you update the figures to (1) show WT variability for each tool instead of for tool pairs (2) remove difference between BT and WT (3) show the XY plot as a joint histogram?}

@@

Table~\ref{table:pipeline-stats} presents summary statistics for variability in the
group-level t-statistics. For each tool pair (A, B), BT variability was
significantly larger than machine error in tool A or B, in both thresholded
and unthresholded maps (Wilcoxon signed-rank test and t-test p~\textless~$10^{-5}$ for all tests).
% \gk{we don't have sufficient sample size to say that p is that small,
% so we should probably say less than $10^{-5}$ or something else we could actually
% arrive at with our number of repetitions, pipeline pairs, and samples.}
% for all tests, t-test p=0 \gk{similar comment re p value} for all tests) \camille{p is smaller than a very small p but not zero?}.
Figure~\ref{fig:unthresh-maps}-\textbf{A},\textbf{B},\textbf{C} shows that
these global differences were confirmed regionally.
In addition, BT and WT variability appeared moderately correlated across voxels (Pearson's r
in [0.60, 0.62], p$< 10^{-5}$, Figure~\ref{fig:unthresh-maps}-\textbf{D}). 
% Interestingly, the correlation appears driven by a set of voxels exhibiting comparable BT and
% WT values located around the identity line in Figure~\ref{fig:unthresh-maps}D, which, however, did not seem to have
% any spatial consistency.

@@

Table~\ref{table:pipeline-stats} also presents summary statistics for
subject-level unthresholded t-statistics maps. As for group-level maps,
BT variability was consistently larger than WT variability (Wilcoxon
signed-rank test and t-test p~\textless~$10^{-5}$ for all tests).
However, Figure~\ref{fig:unthresh-maps-sbj} shows that for
some subjects, WT variability approached %and even surpassed
BT in some regions. As for the group maps, BT and WT
variability appeared moderately correlated for all subjects (r in [0.484,
    0.633], p$< 10^{-5}$ for all subjects, see Supplemental Material~\ref{sec:supp-subjects}).

@@

While the previous results were obtained for machine error, we also
evaluated numerical variability across different virtual precisions for FSL
and found that the virtual precision of t=17 bits minimized the RMSE
between BT and WT in unthresholded group maps.
FSL produced unthresholded group maps with $\mu=0.483$ and $\sigma=0.410$ at this virtual precision,
which was still lower than BT variability
(Wilcoxon signed-rank test and t-test p~\textless~$10^{-5}$ for all tests).
However, Figure~\ref{fig:gnp-mni} shows that
BT and WT reached comparable magnitudes in some regions. BT and WT remained
moderately correlated at this precision \tristan{add r score}.

@@

%%%%%%%%%% Var. of Thresh %%%%%%%%
\begin{figure*}[ht]
  \begin{subfigure}[ht]{\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/plots/wt-tstats/act_deact-framed.png}
    % \includegraphics[width=\textwidth]{figures/abs/gl-thresh.png}
    %\caption{Standard deviation of thresholded t-statistics map on template surface}
  \end{subfigure}
  \centering
  \caption{Thresholded group-level t-statistics within tools at machine error for FSL (\textbf{A}),
    SPM (\textbf{B}), AFNI (\textbf{C}), and confusion matrices of activation instability
    in BT and WT among the 360 regions of the HCP-MMP1.0 parcellation (\textbf{D}).}
  \label{fig:thresh-maps}
\end{figure*}

@@

Figure~\ref{fig:unthresh-worst-sbj} plots the maps of t-statistics for each run of tools in WT.

\begin{figure*}[ht]
  \centering
  \includegraphics[width=\textwidth]{figures/plots/wt-tstats/5-framed.png}
  % \includegraphics[width=.6\textwidth]{figures/sbj-abs-corr-unthresh-plot.png}
  \caption{For subject with highest WT variability, unthresholded subject-level t-statistics
  within tools at machine error for FSL (\textbf{A}),
  SPM (\textbf{B}), AFNI (\textbf{C}).}
  \label{fig:unthresh-worst-sbj}
\end{figure*}

>>>>

As a result of these factors, the variability resulting from the use of
different fMRI analysis tools implementing similar analytical approaches
(between-tool variability) can reach worrying magnitudes. For instance, the
study in~\cite{bowring2019exploring} compared the results produced by the
three main fMRI analysis toolboxes, namely SPM~\cite{penny2011statistical},
AFNI~\cite{cox1996afni} and FSL~\cite{jenkinson2012fsl}, using similar
pipelines. It reported limited similarity between the activation clusters
produced by these tools, measured by Dice coefficients ranging from 0.0 to
0.769, where 0 indicates non-overlapping
clusters and 1 means identical clusters. More recently, the work
in~\cite{Li2021.12.01.470790} also showed a low similarity between the
results produced by five different tools (ABCD~\cite{feczko2021adolescent},
CCS~\cite{xu2015connectome}, CPAC~\cite{craddock2013towards}, DPARSF~\cite{yan2010dparsf},
fMRIPrep~\cite{esteban2019fmriprep}) and identified the main
factors contributing to these differences. The magnitude of the highlighted
differences suggest that between-tool variability may be playing a critical
role in the reproducibility of fMRI analysis overall.

@@

We evaluated BT variability for thresholded as well as unthresholded
group-level and subject-level t-statistics maps by computing the absolute
differences of t-statistic maps across tools. For WT, we computed the
average absolute difference across the three Fuzzy Libmath samples.
Moreover, we computed WT variability for a pair of tools as the sum of WT values for each tool.

@@

Table~\ref{table:pipeline-stats} presents summary statistics for variability in the
group-level t-statistics. For each tool pair (A, B), BT variability was
significantly larger than machine error in tool A or B, in both thresholded
and unthresholded maps (Wilcoxon signed-rank test and t-test p~\textless~$10^{-5}$ for all tests).
% \gk{we don't have sufficient sample size to say that p is that small,
% so we should probably say less than $10^{-5}$ or something else we could actually
% arrive at with our number of repetitions, pipeline pairs, and samples.}
% for all tests, t-test p=0 \gk{similar comment re p value} for all tests) \camille{p is smaller than a very small p but not zero?}.
Figure~\ref{fig:unthresh-maps}-\textbf{A},\textbf{B} shows that
these global differences were confirmed regionally.
In addition, BT and WT variability appeared moderately correlated across voxels (Pearson's r
in [0.60, 0.62], p$< 10^{-5}$, Figure~\ref{fig:unthresh-maps}-\textbf{C}). 
% Interestingly, the correlation appears driven by a set of voxels exhibiting comparable BT and
% WT values located around the identity line in Figure~\ref{fig:unthresh-maps}D, which, however, did not seem to have
% any spatial consistency.

@@

Table~\ref{table:pipeline-stats} also presents summary statistics for
subject-level unthresholded t-statistics maps. As for group-level maps,
BT variability was consistently larger than WT variability (Wilcoxon
signed-rank test and t-test p~\textless~$10^{-5}$ for all tests).
However, Figure~\ref{fig:unthresh-maps-sbj} shows that for
some subjects, WT variability approached %and even surpassed
BT in some regions. As for the group maps, BT and WT
variability appeared moderately correlated for all subjects (r in [0.484,
    0.633], p$< 10^{-5}$ for all subjects, see Supplemental Material~\ref{sec:supp-subjects} and~\ref{sec:supp-worst-subject}).

@@

While the previous results were obtained for machine error, we also
evaluated numerical variability across different virtual precisions for FSL
and found that the virtual precision of t=17 bits minimized the RMSE
between BT and WT in unthresholded group maps.
FSL produced unthresholded group maps with $\mu=0.483$ and $\sigma=0.410$ at this virtual precision,
which was still lower than BT variability
(Wilcoxon signed-rank test and t-test p~\textless~$10^{-5}$ for all tests).
However, Figure~\ref{fig:gnp-mni} shows that
BT and WT reached comparable magnitudes in some regions. BT and WT remained
moderately correlated at this precision (Pearson's r
in [0.53, 0.61], p$< 10^{-5}$, Figure~\ref{fig:gnp-mni}-\textbf{B}).

@@

%%%%%%%%%% Var. of Thresh %%%%%%%%
\begin{figure*}[ht]
  \begin{subfigure}[ht]{\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/act_deact-marked.pdf}
    % \includegraphics[width=\textwidth]{figures/plots/wt-tstats/act_deact-framed.png}
    % \includegraphics[width=\textwidth]{figures/abs/gl-thresh.png}
    %\caption{Standard deviation of thresholded t-statistics map on template surface}
  \end{subfigure}
  \centering
  \caption{Thresholded group-level t-statistics within tools at machine error for FSL (\textbf{A}),
    SPM (\textbf{B}), AFNI (\textbf{C}), and confusion matrices of activation instability
    in BT and WT among the 360 regions of the HCP-MMP1.0 parcellation (\textbf{D}).}
  \label{fig:thresh-maps}
\end{figure*}

@@

Figure~\ref{fig:unthresh-worst-sbj} plots the maps of t-statistics
for each run of tools for the subject with highest WT variability.

\begin{figure*}[ht]
  \centering
  \includegraphics[width=\textwidth]{figures/5-marked.pdf}
  % \includegraphics[width=\textwidth]{figures/plots/wt-tstats/5-framed.png}
  % \includegraphics[width=.6\textwidth]{figures/sbj-abs-corr-unthresh-plot.png}
  \caption{For subject with highest WT variability, unthresholded subject-level t-statistics
  within tools at machine error for FSL (\textbf{A}),
  SPM (\textbf{B}), AFNI (\textbf{C}).}
  \label{fig:unthresh-worst-sbj}
\end{figure*}
